---
title: "Johnson&Johnson ARIMA"
output: 
  html_document:
    keep_md: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### packages
```{r, results="hide", warning=FALSE, message=FALSE}
packages <- c("dplyr", "astsa", "ggfortify", "stats", "fpp2", "tseries", "AICcmodavg", "lmtest", "forecast")

installed_packages <- packages %in% row.names(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed.packages])
}

lapply(packages, library, character.only = TRUE)
``` 

### Import JohnsonJohnson (J&J) quarterly earnings time-series
```{r}
df <- JohnsonJohnson
df
```
### Plot time series
```{r}
tsplot(df)
```

### Lag-plot
```{r}
lag1.plot(df,4)
```
### Seasonal plot
```{r}
ggmonthplot(df)+
  ylab("earnings") +
  ggtitle("Seasonal plot: earnings J&J ")
```
### Seasonal plot over the years
```{r}
ggseasonplot(df)
```
The plots suggest a change in variance over time; an increasing trend; and seasonality.

### Box-Cox transformation to deal with changes in variance
```{r}
lambda <- BoxCox.lambda(df)
print(lambda)
df <- BoxCox(df, lambda = lambda)

plot.ts(df)
```

## De-trend and de-seasonalize

### 1) Fourier transform
```{r}
fourier.df <- tslm(df ~ trend + fourier(df, K=2))
summary(fourier.df)
autoplot(df, series="Data") +
  autolayer(fitted(fourier.df), series="Fitted") +
  xlab("Year") + ylab("Earnings") +
  ggtitle("earnings J&J")
tsplot(residuals(fourier.df), main = "detrended desasonalized beer data")
seasonplot(residuals(fourier.df))
```
### linear transform trend & season
```{r}
fit.df <- tslm(df~ trend + season)
summary(fit.df)
```
### plot adjusted series & residuals
```{r}
autoplot(df, series="Data") +
  autolayer(fitted(fit.df), series="Fitted") +
  xlab("Year") + ylab("Megalitres") +
  ggtitle("Quarterly Beer Production")
tsplot(residuals(fit.df), main = "detrended desasonalized beer data")
```
### Compare AIC of both models
```{r}
models <- list(fourier.df, fit.df)

#calculate AIC of each model
AIC_fourier <- AIC(fourier.df)
AIC_linear <- AIC(fit.df)
cbind(AIC_fourier, AIC_linear)
```
### 3) Loess de-trending
```{r}
df.stlper=stl(df, s.window="periodic")
df.stl=stl(df, s.window=5)
plot(df.stlper)
plot(df.stl)
```

### Inspect the remainder correlation 
```{r}
lag1.plot(df,4)
lag1.plot(df.stl$time.series[,3],4)
```

### 4) Difference de-trend & de-seasonalize
```{r}
#tsplot(diff(df))
tsplot(diff(diff(df)))
df <- diff(diff(df), 4)
```

```{r}
acf1 <- acf(df)
pacf1 <- pacf(df)
```

```{r}
lag1.plot(df,4)
```
Still some correlation remains.

### Dickey-Fuller Test
```{r}
adf.test(df)
```
Suggests stationarity (p-value <0.05).


### Compare with the original ts
```{r}
df_original <- JohnsonJohnson
df_diff <- df

plot.ts(cbind(df_original, df_diff))
```

---End of the assignment---

### Fit ARIMA model
```{r}
#I:2 (d)
#AR:1 (p)
#MA:1 (q)
#ARIMA(pdq) = (1,2,1)
df <- JohnsonJohnson

fitSARIMA <- Arima(df, order=c(1,2,1), seasonal = list(order = c(1,0,0), period = 4),method="ML", lambda=lambda)
coeftest(fitSARIMA)

fitARIMA <- Arima(df, order=c(1,2,1), method="ML",lambda=lambda)
coeftest(fitARIMA)

AIC_SARIMA <- AIC(fitSARIMA)
AIC_ARIMA <- AIC(fitARIMA)
cbind(AIC_ARIMA, AIC_SARIMA)
```
### check residuals
```{r}
tsdisplay(fitSARIMA$residuals)
qqnorm(fitSARIMA$residuals); qqline(fitSARIMA$residuals) 
```
### Plot SARIMA vs df
```{r}
plot(df)
lines(fitted(fitSARIMA), col="red")
```
### Ljung Box-test
```{r}
Box.test(resid(fitSARIMA),type="Ljung",lag=20,fitdf=1)
```

### Forecast
```{r}
Forecastmodel = forecast(fitSARIMA, h = 10, lambda = lambda, level = c(95)) 

plot(Forecastmodel)
```
### train-test split
```{r}
train <- head(df, round(length(df) * 0.95))
h <- length(df) - length(train)
test <- tail(df, h)
autoplot(train) + autolayer(test)
```

### Fit Sarima train-set for predicting next year earnings (4 quarters)
```{r}
fitSARIMA <- Arima(train, order=c(1,2,1), seasonal = list(order = c(1,0,0), period = 4),method="ML", lambda=lambda)
```

### Forecast
```{r}
#Forecastmodel = forecast(train, h = 4, lambda = lambda, level = c(95)) 

fitSARIMA %>%
  forecast(train, h = 4, lambda = lambda, level = c(95)) %>%
  autoplot() + autolayer(test)
```
### Metrics
```{r}
arima.test <- Arima(test, model=fitSARIMA)
accuracy(arima.test)
```
